### 2013-10-08 chi2014/qualitative.tex
new personas

The second category of strategies is \emph{riffing}.
We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: specificity riffing, exhaustive mutation, exploratory riffing, and follow-up riffing. We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.

@@

Finally, \emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented. The new ideas relies on the implications or consequences of the previous. For example:

@@

Within and across these strategies, there are a few repeated phenomena.
Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired exhaustive riffing from a participant:

@@

Almost all participants employ some form of riffing. This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas". However, the different strategies of riffing have very different impacts on the qualitative assessment of a run. Exhaustive mutation is often unsatisfying, as it feels as though the participant is trying to wring every scrap of task completion out of a single idea. On the other hand, exploratory riffing can produce answers that are unexpected and of high quality.

@@

When a solution is provided, it is still occasionally lacking in sufficient information to act upon. We call this strategy \emph{goal without plan}. For example, a common response to the problem was "use as art". In this case, an end goal has been established, but any implementation of the goal requires further idea generation.

@@

We use personas to capture collections of strategy use we see repeated across participants. 
Six personas were identified: quitters, lazy creatives, diligent struggles, lazy strugglers, hard workers, and over-deliverers. These personas are described below. A summary, in terms of strategies used, is in Table TAB.

\emph{Quitters} are the easiest to summarize. These participants do not provide all the requested responses, and often resort to cop-outs or exhaustive mutation in the ideas that they do provide.

\emph{Lazy creatives} provide primarily uncommon responses, but do not put effort into the responses. They employ brevity, reach back and exhaustive mutation strategies to a great extent. These participants seem like prime targets for future interventions, as their ideas are of interest but they focus their strategies towards rapid completion. Lazy creatives employ minimal feature utilization.

\emph{Diligent strugglers} are the opposite of lazy creatives. These participants do not employ the brevity or exhaustive mutation strategies, but nonetheless produce primarily common, unoriginal ideas. These participants also seem like potential targets for interventions, as they apply themselves to the task but may need external assistance to move beyond their comfort zone. Strategies employed include exploratory riffing and feature utilization.

\emph{Lazy strugglers} are simply those who produce no creative ideas and are predisposed towards unproductive strategies (primarily exhaustive mutation and brevity).

\emph{Hard workers} are notable for their minimal use of riffing strategies. The riffing strategies they do use tend to exclude exhaustive mutation. While brevity is occasionally employed for short periods, it is generally only where a short description is sufficient to communicate an idea.

\emph{Over-deliverers} are notable for their lack of strategy in general. Their ideas are verbose, distinct and feature extremely minimal riffing.

\begin{table*}
    \begin{tabular}{|l|p{0.35\linewidth}|l|}
        \hline
        \textbf{Persona} & \textbf{Common strategies} & \textbf{Rare strategies} \\
        \hline
        Quitters & cop-outs, exhaustive mutation & follow-up riffing \\
        \hline
        Lazy creatives & exploratory riffing, specificity riffing, \newline
        exhaustive mutation, brevity, reach-backs  & follow-up riffing  \\
        \hline
        Diligent strugglers & exploratory riffing & exhaustive mutation, brevity, \newline
        follow-up riffing \\
        \hline
        Lazy strugglers & exhaustive mutation, exploratory riffing, \newline
        cop-outs, brevity  & follow-up riffing \\
        \hline
        Hard workers & feature use, humour, \newline
        exploratory riffing, specificity riffing  & follow-up riffing, exhaustive mutation \\
        \hline
        Over-deliverers & feature use & riffing, reach-backs, brevity \\
        \hline
    \end{tabular}
    \caption{Strategies used by personas}
\end{table*}

>>>>

The second category of strategies is \emph{riffing}.
We found riffing manifested itself many ways, and identified four distinct riffing strategies used by participants: generalization riffing, repeat riffing, hold riffing, and continuation riffing. We describe each of these in detail, and then touch on three common phenomena found under riffing strategies.

@@

Hold riffing can also involve holding language terms constant even when the meaning of the term changes between responses, as in free association. For example:

\begin{itemize}
    \item beer coaster
    \item drink coaster (generalization riffing)
    \item roller coaster (hold riffing)
\end{itemize}

\emph{Continuation riffing} is the strategy of creating another idea that assumes the previous idea has already been implemented. The new ideas relies on the implications or consequences of the previous. For example:

@@

Within and across these strategies, we identified two sub-strategies or tweaked behaviours.
Some participants \emph{pair} their riffs. These participants  often riff on old ideas but never produce more than two versions of each idea. For example, paired hold riffing from a participant:

@@

Almost all participants employ some form of riffing. This is not surprising, as it is ingrained in the rules of brainstorming given at the beginning of the task: "combinations of ideas count as new ideas".

@@

When a solution is provided, it is still occasionally lacking in sufficient information to act upon. We call this strategy \emph{goal without plan}. For example, a common response to the problem was "use as art". In this case, an end goal has been established, but any implementation of the goal requires further idea generation. The inverse of this strategy is \emph{plan without goal}, as in the response "change the settings". In this case, one could certainly perform the operation on the MP3 player, but it doesn't provide any obvious benefit.

In many cases, responses that do not have an explicit goal or plan nonetheless imply one. For example, "throw it in the ocean" is an action performed with the MP3 player with no stated goal, but the implied goal is to get rid of the broken device. In these cases, as judged by the coder, reponses were not considered part of a partial solution strategy.

@@

A persona can be thought of as a toolbox of strategies that we see repeated across participants. Participants of a given persona have a subset of strategies that are used often, and a subset that are rarely used. In a process analogous to affinity diagramming, brainstorming runs were clustered based on similarity of strategy use. This resulted in six clusters of repeated behaviour, each of which is a basis of a persona. These personas are summarized in Table TAB.

The \emph{high concept} persona consists of participants who view the MP3 player as an abstract object and provide many distinct, if often vague, responses. Commonly employed strategies are scoping outward, hold riffing, and partial responses.

\emph{Outside the box} personas are similar to high concept, except that their responses are more well-defined. Common strategies are scoping outward, hold riffing and brevity. Partial responses are rare.

and so on...

\subsubsection{Outliers}

In addition to the observed personas, there were several outliers. One participant employed no strategies to a signficant degree (in the case of problem scoping, they scoped inward and outward with similar frequency). A second outlier responded with questions, such as "What can the average person do to better the environment?".

\begin{table*}
    \begin{tabular}{|l|p{0.35\linewidth}|l|}
        \hline
        \textbf{Persona} & \textbf{Common strategies} & \textbf{Rare strategies} \\
        \hline
        High concept & scoping outward, hold riffing, partial responses & brevity, reach-backs \\
        \hline
        Outside the box & scoping outward, hold riffing, brevity & partial responses, reach-backs \\
        \hline
        broad maximizers & repeat riffing, brevity, \newline
        partial responses  & reach-backs \\ 
        \hline
        broad explorers & hold riffing, partial responses & \\
        \hline
        kitchen sink & scoping outward, repeat riffing, brevity, \newline
        partial responses, reach-backs & \\ 
        \hline
        detail explorers & scoping inward, hold riffing & brevity, partial responses, reach-backs\\
        \hline
    \end{tabular}
    \caption{Strategies used by personas}
\end{table*}
